---
title: "Week 03 Assignment 02_datasets"
author: "Payton Bokowy"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    template: ./WSU_STATS419_FALL2020/MonteShaffer/latex-templates/homework.tex
  html_document:
    df_print: paged
course: "STATS 419 Survey of Multivariate Analysis"
course-short: STATS419
instructor: Monte J. Shaffer
email: payton.bokwy@wsu.edu
wsu-id:
fig_caption: true
number_sections: true
---
# Matrix
```{r, matrix}
x = matrix(c(1, 0, 2,
             0, 3, 0,
             4, 0, 5), nrow = 3, byrow = T)

rotateMatrix90 = function(mat)
  {
    t(apply(mat, 2, rev))
  }
rotateMatrix90(x)

rotateMatrix180 = function(mat)
{
  old = t(apply(mat, 2, rev))
  t(apply(old, 2, rev))
}
rotateMatrix180(x)

rotateMatrix270 = function(mat)
{
  apply(mat, 1, rev)
}
rotateMatrix270(x)
```

# Iris Plot
```{r, iris-plot}
library(datasets)
data(iris)
pairs(iris[1:4], main = "Iris Data (red = setosa, green = versicolor, blue = virginica)", pch = 21, bg = c("red", "green3", "blue") [unclass(iris$Species)])
```

# Iris Description
The IRIS data set takes four measurements from three species of iris plants; iris setosa is located on the coast of Alaska, iris versicolor is located in the north eastern portion North America, and iris virginica is located in the north western portion of North America. The sepal length and width and the petal length and width were measured for each sample.

# Personality
```{r, personality-data}
library(lubridate);

personality_raw = read.delim("personality-raw.txt", header = T, sep = "|")

personality = subset(personality_raw, select = -(V00))

new_date = strptime(personality$date_test, format = '%m/%d/%Y %H:%M')
new_personality = cbind(personality, new_date)

new_personality$Year = year(new_date)
new_personality$Week = week(new_date)
new_personality$date_test <- new_personality$new_date <- list(NULL)

sorted_p = new_personality[with(new_personality, order(Year, Week, decreasing = T)), ]

clean = unique(sorted_p);
clean = clean[, -which(names(clean) %in% c("date_test", "new_date"))]

write.table(clean, file = "personality-clean.txt", sep = "|")
```
My raw data had 838 records, and the clean dataset only has 822 records.

# Summary Functions
```{r, summary-functions}
data = read.delim("personality-clean.txt", header = T, sep = "|")
clean = data[, -c(1, 62:63)]
clean_data = data.matrix(head(clean, 1))
d = as.numeric(clean_data)

doSummary = function(x)
{
  length = length(x)
  sum = sum(is.na(x))
  mean = mean(x)
  median = median(x)
  mode = doMode(x);
  myVariance = doSampleVariance(x, "naive")
  sd = sd(x) #compare to custom function
  mySd = sqrt(doSampleVariance(x, "naive"))
  
  result = c(length, sum, mean, median, mode, myVariance, sd, mySd)
  result;
}

doSampleVariance = function(x, method)
{
  sum = 0; sumSq = 0; var = 0; n = 0; i = 0;
  
  if(method == "naive")
  {
    for(i in 1:length(x))
    {
      n = n + 1;
      sum = sum + x[i];
     sumSq = sumSq + x[i]*x[i];
    }
    var = ((sumSq - (sum^2)/n)/(n-1))
    var;
  }
  
  else #two-pass algorithm
  {
    for(i in 1:length(x))
    {
      n = n + 1;
      sum = sum + x[i]
    }
    
    mean = (sum / n)
    
    for (i in 1:length(x))
    {
      sumSq = sumSq + (i - mean) * (i - mean)
    }
    
    var = (sumSq / (n - 1))
  }
  
  result = c(sum, sumSq, var);
  result;
}

doMode = function(x)
{
  #use frequency of data --highest frequency
  test = unique(x)
  freq = tabulate(match(x, test))
  
  result = test[freq == max(freq)]
  result;
}

doSummary(d);
print("My standard deviation and the program's are nearly the same.")
doSampleVariance(d, "naive");
doMode(d);

for(i in 1:length(d))
{
  x[i] = d[i];
  x;
}

Y = scale(x, center = T, scale = T) #calculates z-scores
plot(x, Y, xlab = "Raw Scores", ylab = "Z-Scores", main = "Personality Data")
```
As the raw score goes up, so does the z-score. This makes sense because the lower and higher values would be away from the mean.

# Inflation Dates with Millions
```{r, inflation-rate}
library(devtools); library(stringr); library(rvest);

grabFilmInfoFromFilmsPage = function(page)
	{
	movies = page %>%
		html_nodes(".mode-detail");
		
	pagecount = length(movies);
	
	result = data.frame(matrix(ncol = 11,nrow = pagecount));
	colnames(result) = c("rank", "title", "ttid", "year", "rated", "minutes", "genre", "ratings", "metacritic", "votes", "millions"); 
				
	for(i in 1:pagecount)
		{
		movie = movies[i];
		
		rank = movie %>%
			html_node(".lister-item-index") %>%
			html_text() %>%
			as.numeric();
		result$rank[i] = rank;
			
		title = movie %>%
			html_node(".lister-item-header a") %>%
			html_text();
		result$title[i] = title;
			
		ttid = movie %>%
			html_node(".lister-item-header a") %>%
			html_attr("href");
			
			temp = strsplit(ttid,"/",fixed=T);
		ttid = temp[[1]][3];
		result$ttid[i] = ttid;
		
		year = movie %>%
			html_node(".lister-item-year") %>%
			html_text();
		year = cleanupYear(year);
		result$year[i] = year;
		
		rated = movie %>%
			html_node(".certificate") %>%
			html_text();
		result$rated[i] = rated;
			
		minutes = movie %>%
			html_node(".runtime") %>%
			html_text();
		minutes = cleanupMinutes(minutes);
		result$minutes[i] = minutes;		
		
		genre = movie %>%
			html_node(".genre") %>%
			html_text();
		genre = str_trim(genre);
		result$genre[i] = genre;
		
		ratings = movie %>%
			html_node("div .rating-list") %>%
			html_attr("title");
				temp = strsplit(ratings,"/",fixed=T);
				temp = gsub("Users rated this","",temp[[1]][1],fixed=T);	
				temp = str_trim(temp);
		ratings = as.numeric(temp);
		result$ratings[i] = ratings;
		
		metacritic = movie %>%
			html_node(".ratings-metascore span") %>%
			html_text();
		metacritic = as.numeric(str_trim(metacritic));
		result$metacritic[i] = metacritic;
		
		info = movie %>%
			html_nodes(".lister-item-content p span") %>%
			html_text();
			
		votes = as.numeric(gsub(",","",info[8],fixed=T));
		result$votes[i] = votes;
		
		millions = cleanupMillions(info[11]);
		result$millions[i] = millions;			
		}
		
	result;
	}

cleanupMillions = function(millions)
	{
	millions = gsub('$','',millions, fixed=T);
	millions = gsub('M','',millions, fixed=T);
	
	millions = as.numeric(millions);
	millions;
	}
	
cleanupMinutes = function(minutes)
	{
	minutes = gsub('min','',minutes, fixed=T);
	
	minutes = as.numeric(minutes);
	minutes;
	}
	
cleanupYear = function(year)
	{
	year = gsub('(','',year, fixed=T);
	year = gsub(')','',year, fixed=T);
	year = gsub('I','',year, fixed=T);
	year = as.numeric(year);
	year;
	}

grabNameFromFilmsPage = function(page)
	{
	name = page %>%
		html_node(".header") %>%
		html_text();
		
		name = gsub("Most Rated Feature Films With","",name,fixed=T);
		name = str_trim(name);
	
	name;
	}
	
grabFilmCountFromFilmsPage = function(page)
	{
	totalcount = page %>%
		html_nodes(".desc") %>%
		html_text();
		
		temp = strsplit(totalcount,"of",fixed=T);
		temp2 = strsplit(temp[[1]][2],"titles", fixed=T);
		
		totalcount = str_trim(temp2[[1]][1]);
		totalcount = as.numeric(totalcount);
		
		temp2 = strsplit(temp[[1]][1],"to", fixed=T);
		
		pagecount = str_trim(temp2[[1]][2]);
		pagecount = as.numeric(pagecount);
		
	result = list();
	
	result$totalcount = totalcount;
	result$pagecount = pagecount;
	
	result;
	}

# https://www.imdb.com/title/tt0466839/?ref_=filmo_li_tt ... get box office budget/gross if NA ... on millions. ..

grabFilmsForPerson = function(nmid)
	{
	url = paste("https://www.imdb.com/filmosearch/?explore=title_type&role=",nmid,"&ref_=filmo_ref_typ&sort=num_votes,desc&mode=detail&page=1&title_type=movie", sep="");
	
	page1 = read_html(url);
	result = list();
	## useful for other data purposes
	result$nmid = nmid;
	
	## name of person
	result$name = grabNameFromFilmsPage(page1);
	result$countfilms = grabFilmCountFromFilmsPage(page1);
	
	result$movies.50 = grabFilmInfoFromFilmsPage(page1);
	
	result;
	}
	
infl = "https://www.officialdata.org/us/inflation/2000?endYear=1920&amount=1000000";

infl.html = read_html(infl);

infl.table = infl.html %>%
	html_node(".expand-table-parent") %>%
	html_node(".table-striped") %>%
	html_node("tbody") %>%
	html_nodes("tr");
	
result = data.frame( matrix(nrow=length(infl.table), ncol=3));
colnames(result) = c("year","dollar","inflation");
	
for(i in 1:length(infl.table) )
	{
	infl.row = infl.table[i]	%>% 
		html_nodes("td") %>%
		html_text();
		
	year = as.numeric(infl.row[1]);
		temp = gsub('$','',infl.row[2],fixed=T);
		temp = gsub(',','',temp,fixed=T);
	dollar = as.numeric(temp);
		temp = gsub('%','',infl.row[3],fixed=T);
	inflation = as.numeric(temp);	
	
	result$year[i] = year;
	result$dollar[i] = dollar;
	result$inflation[i] = inflation;
}
  
result[101, 3] = 1.35 #add inflation rate for 2020
new_inflation = NULL;
first = result$dollar[2:101];
second = result$dollar;

for(i in 1:length(second)) #calculates new inflation rate
{
  new_inflation[i] = first[i] /second[i] - 1;
  options(digits = 20);
}

result2 = cbind(result, new_inflation); #combines the new inflation rate in the table
result2['new_inflation'] = c(NA, head(result2['new_inflation'], dim(result2)[1]-1)[[1]]);

result2[1, 4] = .1560999999999999943;

base = result2$dollar[1] / result2$dollar[81]; #sets base as 2000

result2$new_dollar = result2$dollar * base;
new_dollar = formatC(as.numeric(result2$new_dollar), format = "f", digits = 2)
result2$new_dollar = NULL; result2$inflation = NULL;
result3 = cbind(result2, new_dollar);

result3$inflation_rate = base * result3$new_inflation;

nmid = "nm0000226";
will = grabFilmsForPerson(nmid);

nmid = "nm0000243";
denzel = grabFilmsForPerson(nmid);

Will = as.data.frame(will$movies.50);
Denzel = as.data.frame(denzel$movies.50);

willOrdered = Will[with(Will, order(year)),];
denzelOrdered = Denzel[with(Denzel, order(year)),];

inflationRates = result3[-c(1:73),];

WillTable = merge(willOrdered, inflationRates, by = c("year"))
DenzelTable = merge(denzelOrdered, inflationRates, by = c("year"));

WillTable$millions.2000 = WillTable$millions * WillTable$inflation_rate;

DenzelTable$millions.2000 = DenzelTable$millions * DenzelTable$inflation_rate;

par(mfrow=c(1,2));
boxplot(WillTable$millions.2000, main = will$name, ylim=c(0,1.2), ylab = "Adjusted Millions");
boxplot(DenzelTable$millions.2000, main = denzel$name, ylim=c(0,1.2), ylab = "Adjusted Millions");
par(mfrow=c(1,1));
```

# Compare Will Smith & Denzel Washington
```{r, comparisons}
par(mfrow=c(1,2));
boxplot(WillTable$millions.2000, main = will$name, ylim=c(0,1.2), ylab = "Adjusted Millions");
boxplot(DenzelTable$millions.2000, main = denzel$name, ylim=c(0,1.2), ylab = "Adjusted Millions");
par(mfrow=c(1,1));
```
Will Smith has more movies above the median than Denzel Washington. Although, most of Smith’s movies have many movies that made a significant amount of money, both have similar median values for money their movies had made.

```{r}
par(mfrow=c(1,2));
boxplot(WillTable$votes, main = will$name, ylim=c(0,550000), ylab = "Total Votes per Movie");
boxplot(DenzelTable$votes, main = denzel$name, ylim=c(0,550000), ylab = "Total Votes per Movie");
par(mfrow=c(1,1));
```
Will Smith’s movies have gotten significantly more votes than Denzel Washington’s movies. Washington’s movies have slightly a lower median rating than Smith’s.

```{r}
par(mfrow=c(1,2));
boxplot(WillTable$ratings, main = will$name, ylim=c(0,10), ylab = "Ratings");
boxplot(DenzelTable$ratings, main = denzel$name, ylim=c(0,10), ylab = "Ratings");
par(mfrow=c(1,1))
```
Denzel Washington’s median rating is higher than Will Smith’s by about half a point. Will Smith has more outliers towards the minimum; whereas; Washington’s movies have a smaller range.
